Efficient Methods for Solving String Constraints with Bounded Repetition

### 2023-04-10 2023-1-19/efficient_way_for_counting.tex
atva2023 first version

\usepackage{subcaption}
\usepackage{tikz}
\usetikzlibrary {arrows.meta, automata, positioning, fit}
\usepackage{url}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage[]{float}
% set algorithmicx
\renewcommand{\algorithmicrequire}{\textbf{Input:}}  % Use Input in the format of Algorithm  
\renewcommand{\algorithmicensure}{\textbf{Output:}} % Use Output in the format of Algorithm

@@

\newcommand*{\myvec}[1]{\vec{#1}}
\newcommand*{\regex}{\mathcal{R}}
\newcommand*{\strsort}{\verb|Str|}
\newcommand*{\intsort}{\verb|Int|}
\newcommand*{\lan}{\mathcal{L}}
\newcommand*{\highlight}[1]{\textbf{\textit{#1}}}
\newcommand*{\aut}{\mathcal{A}}
\newcommand*{\algfun}[1]{\texttt{#1}}
\newcommand*{\strlen}[1]{\texttt{len}(#1)}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Efficient Methods for Solving String Constraints with Large Repetition Bound}

@@

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  Regular expression with large repetition bounds frequently appears in the real world. However, state-of-art string solvers can not efficiently solve string constraints with such regular expression, which slows down string solver applications like formal verification and symbolic execution. In this paper, we
  propose a new systematic algorithm based on cost-enriched automaton. We use
  heuristic ways like over-approximation and under-approximation to accelerate the search. Further, we extract amounts of regex expressions with large repetition
  bounds from real-world programs and generate a significant benchmark. We evaluate the algorithm's performance on our developed benchmarks and other typical benchmarks. The result shows that our solver outperforms the state-of-art string solvers.

\keywords{string constraints, cost-enriched automata, regex expression, bounded repetition}
\end{abstract}

\section{Introduction}
Regular expressions are widely used in programming languages. About 30–40 \% of Java, JavaScript, and Python software uses regex matching \cite{redos_fse2019}. To handle the problems in software verification reasoning about string, the string theory is introduced and amounts of string solvers are developed.
However, the string theory is one of the most challenging theories because it is easy to be undecidable \cite{undecidable_2016}.
String solvers try very hard to improve efficiency:  DPLL(T)-based string solver Z3 \cite{z3} and Cvc5 \cite{cvc5} apply many heuristic derivation rules. Automaton-based string solver Trau+\cite{trau} proposed flat automaton and uses the CEGAR framework to under-approximate
and over-approximate the string constraints again and again.
Unfortunately, the string solvers above perform poorly when the string constraints contain large repetition bounds. [add a example]. To address this issue, we extend cost-enriched finite automaton(CEFA)\cite{atva2020} with linear arithmetical constraint. Use backward-propagating algorithm to search a solution. Our main contributions are as follows: 
\begin{enumerate}
  \item  Extend cost-enriched finite automaton to model the regular expression with large repetition bound.
  \item  Devise an efficient algorithm based on CEFA to solve string constraints with large repetition bound.
  \item  Generate many instances with large repetition bound from real-word regular expressions.
  \item  Implemented our efficient algorithm on Ostrich\ref{Ostrich} and compared it with state-of-art string solvers. The result shows the superiority of our automaton model and algorithm.
\end{enumerate}
%TODO: list the main contributions
% The structure of this paper is as follows: We introduce some abbreviations and concepts in section \ref{sec:pre} and introduce the syntax and semantics of the string theory logic in section \ref{sec:str_t}. Then we develop cost-enriched finite automaton model and its operations at section \ref{sec:cefa}. The efficient algorithm is discussed in section \ref{sec:solve}, and the implementation of the algorithm is mentioned in section \ref{sec:imp}. Finally, we conclude our work and look into future work in section \ref{sec:conclu}.

\section{Preliminaries} \label{sec:pre}
%TODO: preliminaries
\highlight{Tokens.} A finite \emph{alphabet} $\Sigma$ is the set of all \emph{letters}. A
\emph{string} is a finite sequence of letters from $\Sigma$. We use $\Sigma^*$ to
denote the set of strings over $\Sigma$, $\epsilon$ to denote the empty string, and $a,
  b,\cdots$ to denote constant letters in $\Sigma$. We use $u, v,\cdots$ to denote constant
string and $x, y,\cdots$ to denote variable string. Moreover, we also consider a set of natural integer numbers $\mathbb{N} = \{0,1,2,3,\cdots\}$ and a set of integer numbers $\mathbb{Z}= \{\cdots, -1, 0, 1, \cdots\}$. We use $m,n,\cdots$ to
denote integer constant, and $i,j\cdots$ to denote integer variable. For vector, we use $\myvec{v}$ to denote the vector of integer constant, $0_n$ to denote the zero vector with length $n$, $v(i)$ to denote the integer value of $\myvec{v}$ at position $i$, $\myvec{v_1}\cdot\myvec{v_2}$ to denote the concatenation of $\myvec{v_1}$ and $\myvec{v_2}$, $R$ to denote the vector of integer variable, $R_1 \cap R_2$ to denote the same variables of $R_1$ and $R_2$, and $|\myvec{v}|$ or $|R|$ to denote the length of vector.\newline
\highlight{CNF, DNF, clause, cube.} We assume that the reader is familiar with first-order logic. A \emph{literal} is an atomic proposition or its negation. A \emph{clause} is a disjunction of literals. The empty clause is $true$. A formula is in \emph{conjunctive normal norm} (CNF) if it is a conjunction of clauses. A \emph{cube} is a conjunction of a consistent set of literals; The empty cube is $false$. A formula is in \emph{disjunctive normal form} (DNF) if it is a disjunction of cubes.\newline
\highlight{Basic Regular Language.} A nondeterministic finite state automaton (NFA) is a tuple $\aut=(Q, \Sigma, \delta, I, F)$ where $Q$ is a finite set of states; $\Sigma$ is a finite alphabet; $\delta\in Q\times\Sigma\times Q$ is the transition relation; $I, F\subseteq Q$ are the set of initial states and finite states respectively. We write a transition $(q, a, q') \in \delta$ as $q\xrightarrow{a} q'$ for readability. A \emph{run} of an NFA $\aut$ on a string $w = a_1\cdots a_n$ is a sequence of transitions $q_0\xrightarrow{a_1}q_1\cdots q_{n-1}\xrightarrow{a_n}q_n$ with $q_0\in I$. The run is \emph{accepting} if $q_{n}\in F$. A word $w$ is \emph{accepted} by an NFA $\aut$ if there is an accepting run on $w$. The \emph{language} of $\aut$ is the set of all words accepted by $\aut$, denoted by $\lan(\aut)$. From automata theory\cite{aut_hopcraft}, basic regular expression only contains union, concatenation, and closure operations. We know that the language of basic regular expression are also in the class of $\lan(\aut)$. We call it \emph{Basic Regular Language}.
\section{String Theory with Regex} \label{sec:str_t}
% TODO: Syntax and semantic
In this section, the syntax and semantic of string theory is defined.
\subsection{Syntax}
In this paper, we propose a new logic called \textit{Extended String Logic (ESL)}. ESL contains the following
sorts: string sort \verb|Str|, integer sort \verb|Int|, the enumerable set of string
sort $\verb|Str|^m$, and the enumerable set of integer sort $\verb|Int|^m$ for each $m\in
  \mathbb{N}$. Furthermore, ESL includes the following predicate and function constants:
string concatenation $\verb|con|:\verb|Str|\times\verb|Str|\rightarrow \verb|Str|$;
string length $\verb|len|: \verb|Str|\rightarrow \verb|Int|$; regular membership $\in:
\verb|Str|\times \verb|Str|^m$; the regex operators such as closure, concatenation,  conjunction, disjunction, intersection, repetition and complement; and the usual constants of linear arithmetic.

The syntax of ESL is presented in Table \ref{tab:syntax}. $\varphi$ is a first-order logic formula that can be a regular membership $x\in\regex$,
a word equation $e$, a quantifier-free Presburger arithmetic formula $\beta$, a negation of a formula, or a
disjunction of two formulas. A word equation $e$ is an equality of two string terms.$\beta$ is an (in)equality of two integer terms and the (in)equality operator $\odot$ contains $=, \geq$. A string term $s$ is an empty string $\epsilon$, letters
$a\in\Sigma$, string variable $x$, or a concatenation of two string terms. An integer
term $\alpha$ is an integer constant 0, integer constant 1, integer variable $i$, the
length of the string term $s$, minus an integer term, or plus of two integer terms. The
regular expression $\regex$ is built on empty string $\epsilon$, constant letter $a\in \Sigma$. The supported regex operations involve concatenation $\cdot$, disjunction $+$,
intersection $\times$, closure $*$, complement $C$, and repetition $\{m, n\}$.

We consider \emph{straight-line} fragment\cite{atva2020}\cite{quadratic_aplas2018} $ESL_{s}$ of the string theory logic. 
A formula $\varphi$ is said to be straight-line if it can be converted to a DNF formula $\bigvee_{i} \phi_i$ and $\phi_i= (\bigwedge_{i=1}^n e_i) \wedge \Re \wedge P$
where $\bigwedge_{i=1}^n e_i$ is straight-line, $\Re$ is a
\emph{cube} of regular membership $x\in \regex$ or its negation $x\not\in \regex$, and $P$ is a Presburger arithmetic formula. A word equation \emph{cube} $\bigwedge_{i=1}^n e_i$ is said to be straight-line if it can be rewritten as the form $\bigwedge_{j=1}^m x_j = s_j$ such that:
\begin{enumerate}
  \item $x_1, \cdots, x_m$ are different variables,
  \item $FV(s_j)\cap \{x_1, x_2, \cdots, x_{j-1}\} = \emptyset$ where $FV(\varphi)$ are all free variables of $\varphi$. \label{straigth_line_condition_2}
\end{enumerate} 
We call straight-line formula $(\bigwedge_{i=1}^n e_i) \wedge \Re \wedge P$ \emph{normal form}.We use $s_1s_2$ to stand for $\verb|concat|(s_1,s_2)$, $|x|$ to
stand for $\verb|len|(x)$.
\begin{table}[H]
  \centering
  \begin{tabular}{l}
    $\regex$ ::= $\epsilon\mid a\mid \regex\cdot\regex\mid \regex+\regex\mid \regex\times \regex\mid \regex^*\mid\regex^C \mid \regex\{m,n\}$ \\
    $\varphi$ ::= $x\in \regex\mid e\mid\beta\mid \neg\varphi\mid \varphi\vee\varphi$                                                      \\
    $e$ ::= $s_1=s_2 $                                                                                                                     \\
    $\beta$ ::= $\alpha_1\odot \alpha_2$                                                                                                   \\
    $s$ ::= $\emptyset \mid \epsilon \mid a\mid x\mid \verb|concat|(s_1, s_2)$                                                             \\
    $\alpha$ ::= $0\mid 1\mid i \mid  \verb|len|(s)\mid -\alpha\mid \alpha_1+\alpha_2$ 
  \end{tabular}
  \caption{Syntax of ESL}
  \label{tab:syntax}
\end{table}

\begin{example}[Straight-line Formula]
  \begin{equation} \label{eq:1}
    \varphi_s ::= x =yz \wedge y = z \wedge x\in a\{1,100\} \wedge |x| = 1
  \end{equation}
  $\varphi_s$ is straight-line because we can split it to three parts: (i) a straight-line word equation cube $x =yz \wedge y = z$,
  (ii) a regular membership cube $x\in a\{1,100\}$, and (iii) a Presburger arithmetic formula $|x| = 1$.
\end{example}

\begin{example}[Non Straight-line Formula]
  \begin{equation} \label{eq:1}
    \varphi_{ns} ::=  x =yz \wedge y = xz \wedge x\in a\{1,100\} \wedge |x| = 1
  \end{equation}
  $\varphi_{ns}$ is not straight-line because the word equation cube\newline
  $x = yz \wedge y = xz$ does not satisfy condition \ref{straigth_line_condition_2}: $FV(yz) \cap \{y\} \not = \emptyset$ and $FV(xz) \cap \{x\} \not = \emptyset$.
\end{example}

\subsection{Semantics}
\highlight{Regular Language} We have introduced basic regular language at
section \ref{sec:pre}. In addition to basic regular operations, we syntactically support intersection, complement and repetition. As we all know, basic regular language is closed under intersection and complement. Furthermore, the operation \emph{repetition} $\regex\{m,n\}$ means repeating regex $\regex$ at least $m$ times and at most $n$ times. It can be syntactically rewritten by concatenation and union:
$\regex\{m,n\} \equiv \regex^m\mid\cdots\mid\regex^n$ where $\regex^k$ defines concatenate $k$ times($m\leq k\leq n$) to $\regex$. So the regular language defined in Table \ref{tab:syntax} is semantically equal to basic regular language. However, if we naively rewrite the repetition operation, the automaton size will linear to the repetition times, causing the search space to be exponential. To address the issue, we proposed a new automaton model called \emph{Cost-Enriched Finite Automaton (CEFA)}, whose size is not relevant to the repetition times. CEFA can not handle nested repetition (e.g., $\regex\{m_1,n_1\} \regex\{m_2,n_2\}$ or $(\regex\{m, n\}) *$) and complementation of repetition (e.g, $(\regex\{m,n\})^C$). So we have to rewrite these regexes syntactically (e.g., $a\{1,100\}\{2,2\}$ is rewritten to $a\{1,100\}a\{1,100\}$).\newline
\highlight{Semantics} We assume that $S$ is the set of string variables over $\Sigma^*$, and $I$ is the set of integer variables. $\eta: S\times\Sigma\rightarrow\Sigma^*$ is the interpretation on string where $\eta(c)=c$ for every letter $c\in \Sigma$ and $\eta(s_1s_2)=\eta(s_1)\eta(s_2)$. $\theta: I\rightarrow\mathbb{Z}$ is the interpretation of arithmetic that is the same as Presburger arithmetic. Then the semantics is given by a satisfaction relation: $\eta, \theta\models \varphi$ defined in Table \ref{tab:semantics}. We say a formula $\varphi$ is \emph{satisfiable} if a solution $(\eta, \theta)$ exists such as $\eta, \theta\models \varphi$. A formula $\varphi$ is \emph{unsatisfiable} if no solution exists.

\begin{table}[H]
  \centering
  \begin{tabular}{lcl}
     $\eta,\theta \models \varphi_1\vee \varphi_2$ & $\mathsf{iff}$ & $\eta,\theta \models \varphi_1 \text{ or } \eta,\theta \models \varphi_2$\\
     $\eta,\theta \models \neg\varphi $          & $\mathsf{iff}$ & $\eta,\theta \not\models \varphi$\\
     $\eta, \theta \models x\in \regex$            & $\mathsf{iff}$ & $\exists w \in \lan(\regex),\eta,\theta \models x = w$\\
     $\eta, \theta \models s_1 = s_2$              & $\mathsf{iff}$ & $\eta(s_1) = \eta(s_2)$\\
     $\eta, \theta \models \alpha_1\odot\alpha_2$  & $\mathsf{iff}$ & $\theta(\alpha_1) \odot \theta(\alpha_2) \text{ where } \odot = \{=, \geq \}$
  \end{tabular}
  \caption{Semantics}
  \label{tab:semantics}
\end{table}
\subsection{Problem}
In this paper, we consider the following problem:
\begin{table}[H]
  \centering
  \begin{tabular}{l l}
    \hline
    \textbf{Problem:} & Is an $ESL_s$ formula $\varphi$ satisfiable?                   \\
    \textbf{Input:}   & A straight-line $ESL_s$ formula $\varphi$ in the normal form. \\
    \textbf{Output:}  & \emph{sat} or \emph{unsat}.                               \\
    \hline
  \end{tabular}
\end{table}

\section{Cost-Enriched Finite Automaton and Basic Operation} \label{sec:cefa}
%TODO: The definition of cost-enriched automata
%      The operation on the cost-enriched automata (union, intersection, complement, concatenate)
In this section, the cost-enriched finite automaton and operations on it is defined. The operations contains intersection, union, complement, concatenation, and repetition.
\subsection{Cost-Enriched Finite Automaton}
\begin{definition}[Cost-Enriched Finite Automaton]
  A cost-enriched finite automaton $\aut$ is a tuple $(Q, \Sigma, \delta, I, F, R, \theta)$ where
  \begin{itemize}
    \item $Q,\Sigma,I,F$ is defined as NFA,
    \item $R = (r_1\cdots r_n)$ is a vector of mutually distinct cost registers,
    \item $\delta$ is the transitions set whose elements are transition tuple $(q, a, q', \myvec{v})$ where $q, q'$ are states in $Q$, $a$ is a letter in alphabet $\Sigma$ and $\myvec{v}$ is vector of natural number. The transition $(q, a, q', \myvec{v})$ is written as $q\xrightarrow[\myvec{v}]{a} q'$ for readability.
    \item $\theta$ is the linear integer arithmetic constraint of registers.
  \end{itemize}
  A \emph{run} of $\aut$ on string $a_1\cdots a_m$ is a transition sequence $q_0\xrightarrow[\myvec{v_1}]{a_1}q_1\cdots q_{m-1}\xrightarrow[\myvec{v_m}]{a_m}q_m$ such that $q_0\in I$ and $\mathcal{V}(r_i) = \displaystyle\sum_{j=1}^m v_j(i)$ is the \emph{value} of register $r_i$ after the run. Note that $\mathcal{V}(r_i)$ is initiated to 0 at state $q_0$. The run is \emph{accepting} if $q_m\in F$ and $\theta[r_i/\mathcal{V}(r_i)]$ is satisfiable. Note that $\theta[r_i/\mathcal{V}(r_i)]$ means a substitution of all occurrences of $r_i$ in $\theta$ to its current value. A string word $w$ is accepted by $\aut$ if there is an accepting run of $\aut$ on $w$. The language of $\aut$ is the set of string words accepted by $\aut$, denoted by $\lan(\aut)$. 
\end{definition}
\begin{example}[The CEFA  $\aut_{(ab)\{1,100\}}$] \label{eg:1}
   Table \ref{fig:repetition} is the CEFA of the regular expression $(ab)\{1,100\}$. The transition set are $\{q_0\xrightarrow[1]{a} q_1, q_1 \xrightarrow[0]{b} q_2, q_2 \xrightarrow[0]{\epsilon} q_0\}$. The register is $r_1$. And the linear arithmetic is $1\leq r_1\leq 100$. Intuitively, each accepted run of $\aut_{(ab)\{1,100\}}$ repeatedly run word $ab$ from one time to 100 times because $1\leq r_1 \leq 100$ restrict the value of $r_1$ to the range $[1,100]$. More precisely, the repeat time of the transition $q_0\xrightarrow[1]{a}q_1$ is restricted to $[1,100]$.
  \begin{figure}[h]
    \centering
    \begin{tikzpicture}[shorten >=1pt,node distance=2cm,on grid,>={Stealth[round]},
      every state/.style={draw=blue!50,very thick,fill=blue!20}, initial text=]

      \node[state,initial]            (q_0)                      {$q_0$};
      \node[state]                    (q_1) [right=of q_0]       {$q_1$};
      \node[state,accepting]          (q_2) [right=of q_1]       {$q_2$};

      \path[->] (q_0) edge              node      [above]           {a/1} (q_1)
      (q_1) edge              node      [above]           {b/0} (q_2)
      (q_2) edge [bend left]  node      [below]           {$\epsilon$/0} (q_0);
    \end{tikzpicture}
    \caption{$\aut_{(ab)\{1,100\}}$ with one register $r_1$ and $\theta = 1\leq r_1 \leq 100$}
    \label{fig:repetition}
  \end{figure}
\end{example}
The example \ref{eg:1} gives a general picture of constructing a CEFA for a regular expression: firstly, we build an NFA from the regex $\regex$
(e.g, To construct CEFA in example \ref{eg:1}, an NFA is construted with
transitions \{$q_0\xrightarrow{a} q_1\xrightarrow{b}q_2\xrightarrow{\epsilon}
  q_0\}$ initial state $q_0$, and accepting state $q_2$). Secondly, we extend NFA to CEFA by adding registers to $R$ and vector to each transition (e.g., the
register $r_1$ is added to store repetition times and add vector $\myvec{1}$ is set to transition $q_0\xrightarrow{a}q_1$ to update the repetition time). Finally, we set the
linear arithmetic $\theta$ carefully to restrict the accepted word to be
included in $\lan(\regex)$ (e.g., $\theta = 1\leq r_1\geq 100$ is set to
restrict the repetition times to $[1,100]$). Besides the repetition operation, other operations are non-trivial on CEFA. We individually discuss them in the following subsections.

\subsection{Non-nested Repetition}\label{subsec:rep}
Given a CEFA $\aut = (Q,\Sigma,\delta,I,F,\emptyset,true)$, the repetition operation $\{m,n\}$ on it is defined as a CEFA $\aut_{rep} = (Q, \Sigma, \delta', I, F, r_1, \theta')$ where: 
\begin{itemize}
  \item $r_1$ is the new register,
  \item $\delta'$ is composed by transitions $q\xrightarrow[\myvec{v}\cdot 0]{a} q'$ for all transitions $q\xrightarrow[\myvec{v}]{a} q' \in \delta$, and transitions $q_m\xrightarrow[0_{n}\cdot 1]{\epsilon} q_0$  for all $q_m\in F$ and $q_0\in I$,
  \item $\theta'$ is the linear arithmetic $m\leq r_1\leq n$ if all initial states are not accepting. Otherwise, $\theta'$ is the linear arithmetic $r_1\leq n$. 
\end{itemize} 
In the construction of $\aut_{rep}$, a new register $r_1$ is added to store the repetition times. The transition $q_m\xrightarrow[0_{n}\cdot 1]{\epsilon} q_0$ is added to update the repetition times. The linear arithmetic $\theta'$ is set to restrict the repetition times to $[m,n]$. Because arbitrary repetition times of empty string is still empty string, the value of $r_i$ need not be greater than $m$ when some initial states are accepting.\newline
Note that the given CEFA $\aut$ does not contain register. So the repetition operation on $\aut$ is non-nested.

\subsection{Concatenation}\label{subsec:con}
Given two CEFA $\aut_1 = (Q_1, \Sigma, \delta_1, I_1, F_1, R_1, \theta_1)$ and $\aut_2 = (Q_2, \Sigma, \delta_2,$ $I_2, F_2, R_2, \theta_2)$ with $R_1\cap R_2= \emptyset$ and $|R_1|=m,|R_2|=n$, the concatenation is defined as $\aut_{con}=(Q_1\cup Q_2, \Sigma, \delta', I_1, F_2, R_1\cdot R_2, \theta_1\wedge\theta_2)$ where $\delta'$ is composed by
\begin{itemize}
  \item $q_1\xrightarrow[\myvec{v_1}\cdot 0_n]{a} q_1'$ for each transition $q_1\xrightarrow[\myvec{v_1}]{a} q_1' \in \delta_1$,
  \item $q_2\xrightarrow[0_m\cdot\myvec{v_2}]{a} q_2'$ for each transition $q_2\xrightarrow[\myvec{v_2}]{a} q_2' \in \delta_2$,
  \item and $q_1\xrightarrow[0_{m+n}]{\epsilon} q_2$ for each $q_1\in F_1$ and $q_2 \in I_2$.
\end{itemize}
% \begin{figure}[h]
%   \centering
%   \begin{subfigure}{0.20\textwidth}
%     \begin{tikzpicture}[shorten >=1pt,node distance=2cm,on grid,>={Stealth[round]},
%       every fit/.style={draw,minimum height=2cm,minimum width=3.5cm,dashed},
%       every state/.style={draw=blue!50,minimum size = 0.1cm}, initial text=]
%       \node[state, initial]   (q_0) {};
%       \node[state, accepting]    (q_1) [right=of q_0] {};
%       \path[->] (q_0) edge [bend right] node [below] {$a/\myvec{v_1}$} (q_1);
%       \node [fit=(q_0) (q_1)] {$\cdots$};
%     \end{tikzpicture}
%     \caption{The CEFA $\aut_1$}
%   \end{subfigure}
%   \begin{subfigure}{0.20\textwidth}
%     \begin{tikzpicture}[shorten >=1pt,node distance=2cm,on grid,>={Stealth[round]},
%       every fit/.style={draw,minimum height=2cm,minimum width=3.5cm,dashed},
%       every state/.style={draw=blue!50,minimum size = 0.1cm}, initial text=]
%       \node[state, initial]   (q_0) {};
%       \node[state, accepting]    (q_1) [right=of q_0] {};
%       \path[->] (q_0) edge [bend right] node [below] {$a/\myvec{v_2}$} (q_1);
%       \node [fit=(q_0) (q_1)] {$\cdots$};
%     \end{tikzpicture}
%     \caption{The CEFA $\aut_2$}
%   \end{subfigure}
%   \begin{subfigure}{0.4\textwidth}
%     \begin{tikzpicture}[shorten >=1pt,node distance=2cm,on grid,>={Stealth[round]},
%       every fit/.style={draw,minimum height=2cm,minimum width=3.5cm,dashed},
%       every state/.style={draw=blue!50,minimum size = 0.1cm}, initial text=]
%       \node[state, initial]   (q_0) {};
%       \node[state]    (q_1) [right=of q_0] {};
%       \node[state]    (q_2) [right=of q_1] {};
%       \node[state, accepting]    (q_3) [right=of q_2] {};
%       \path[->]
%       (q_0) edge [bend right] node [below] {$a/\myvec{v_1}\cdot \red{0_n}$} (q_1)
%       (q_1) edge node [below] {$\red{0_{m+n}}$} (q_2)
%       (q_2) edge [bend right] node [below] {$a/\red{0_m}\cdot \myvec{v_2}$} (q_3);
%       \node [fit=(q_0) (q_1)] {$\cdots$};
%       \node [fit=(q_2) (q_3)] {$\cdots$};
%     \end{tikzpicture}
%     \caption{The concatenation of $\aut_1$ and $\aut_2$}
%   \end{subfigure}
%   \caption{Concatenation}
%   \label{fig:con}
% \end{figure}
The concatenation of two CEFAs is similar to the concatenation of two NFAs, except that the registers and its updates are also concatenated.

% Figure \ref{fig:con} outlines the vector change on transitions when concatenating two CEFA. Without losing information, $\aut_{con}$ contains all registers in $\aut_1$ and $\aut_2$. Furthermore, $\aut_{con}$ update registers' value of $\aut_1$ and $\aut_2$ separately: $\aut_{con}$ only update registers of $R_1$ on the transitions of $\aut_1$ and update registers of $R_2$ on the transition of $\aut_2$. $\theta'$ is the conjunction of $\theta_1$ and $\theta_2$ to make sure all linear constraints in $\aut_1$ and $\aut_2$ are satisfiable.
\subsection{Intersection}\label{subsec:inter}
Given two CEFA $\aut_1 = (Q_1, \Sigma, \delta_1, I_1, F_1, R_1, \theta_1)$ and $\aut_2 = (Q_2, \Sigma, \delta_2,$ $I_2, F_2, R_2, \theta_2)$ with $R_1\cap R_2 = \emptyset$, the intersection is defined as $\aut_{inter} = (Q_1\times Q_2, \Sigma, \delta', I_1\times I_2, F_1\times F_2, R_1\cdot R_2, \theta_1\wedge \theta_2)$ where $\delta'$ is composed by
\begin{itemize}
  \item transitions $(q_1,q_2)\xrightarrow[\myvec{v_1}\cdot\myvec{v_2}]{a} (q_1',q_2')$ for  transition $q_1\xrightarrow[\myvec{v_1}]{a}q_1'$ and transition $q_2\xrightarrow[\myvec{v_2}]{a}q_2'$ exist individually in  $\delta_1$ and $\delta_2$.
\end{itemize}
% \begin{figure}[h]
%   \centering
%   \begin{subfigure}{0.20\textwidth}
%     \begin{tikzpicture}[shorten >=1pt,node distance=2cm,on grid,>={Stealth[round]},
%       every fit/.style={draw,minimum height=2cm,minimum width=3.5cm,dashed},
%       every state/.style={draw=blue!50,minimum size = 0.1cm}, initial text=]
%       \node[state, initial]   (q_0) {};
%       \node[state, accepting]    (q_1) [right=of q_0] {};
%       \path[->]
%       (q_0) edge [bend right] node [below] {$a$/$\myvec{v_1}$} (q_1)
%       (q_0) edge [bend left] node [above] {$b$/$\myvec{v_1}$} (q_1);
%       \node [fit=(q_0) (q_1)] {$\cdots$};
%     \end{tikzpicture}
%     \caption{The CEFA $\aut_1$}
%   \end{subfigure}
%   \begin{subfigure}{0.20\textwidth}
%     \begin{tikzpicture}[shorten >=1pt,node distance=2cm,on grid,>={Stealth[round]},
%       every fit/.style={draw,minimum height=2cm,minimum width=3.5cm,dashed},
%       every state/.style={draw=blue!50,minimum size = 0.1cm}, initial text=]
%       \node[state, initial]   (q_0) {};
%       \node[state, accepting]    (q_1) [right=of q_0] {};
%       \path[->]
%       (q_0) edge [bend right] node [below] {$a$/$\myvec{v_2}$} (q_1)
%       (q_0) edge [bend left] node [above] {$c$/$\myvec{v_2}$} (q_1);
%       \node [fit=(q_0) (q_1)] {$\cdots$};
%     \end{tikzpicture}
%     \caption{The CEFA $\aut_2$}
%   \end{subfigure}
%   \begin{subfigure}{0.4\textwidth}
%     \centering
%     \begin{tikzpicture}[shorten >=1pt,node distance=2cm,on grid,>={Stealth[round]},
%       every fit/.style={draw,minimum height=2cm,minimum width=3.5cm,dashed},
%       every state/.style={draw=blue!50,minimum size = 0.1cm}, initial text=]
%       \node[state, initial]   (q_0) {};
%       \node[state, accepting]    (q_3) [right=of q_0] {};
%       \path[->]
%       (q_0) edge [bend right] node [below] {$a$/\red{$\myvec{v_1}\cdot \myvec{v_2}$}} (q_3);
%       \node [fit=(q_0) (q_3)] {$\cdots$};
%     \end{tikzpicture}
%     \caption{The intersection of $\aut_1$ and $\aut_2$}
%   \end{subfigure}
%   \caption{Intersection}
%   \label{fig:inter}
% \end{figure}
The intersection of two CEFAs is similar to the intersection of NFA, except that the registers updates and linear arithmetic are also intersected.

\subsection{Union}\label{subsec:union}
Given two CEFA $\aut_1 = (Q_1, \Sigma, \delta_1, I_1, F_1, R_1, \theta_1)$ and $\aut_2 = (Q_2, \Sigma, \delta_2,$ $I_2, F_2, R_2, \theta_2)$ with $R_1\cap R_2 = \emptyset$ and $|R_1|=m,|R_2|=n$, the union is defined as $\aut_{union} = (Q_1\cup Q_2\cup\{q_0\}, \Sigma, \delta', \{q_0\}, F_1\cup F_2, R_1\cdot R_2\cdot (r_1, r_2), \theta')$ where $\delta'$ is composed by
\begin{itemize}
  \item transitions $q_0\xrightarrow[0_{m+n}\cdot(1,0)]{\epsilon}q_1$ for all $q_1\in I_1$,
  \item transitions $q_0\xrightarrow[0_{m+n}\cdot(0,1)]{\epsilon}q_2$ for all $q_2\in I_2$,
  \item transitions $q_1\xrightarrow[\myvec{v_1}0_{n+2}]{a} q_1'$ for all $q_1\xrightarrow[\myvec{v_1}]{a} q_1'\in \delta_1$,
  \item transitions $q_2\xrightarrow[0_m\myvec{v_2}0_2]{a} q_2'$ for all $q_2\xrightarrow[\myvec{v_2}]{a} q_2'\in \delta_2$.
\end{itemize}
$r_1$ and $r_2$ are new registers. $q_0$ is a new state where $q_0\not\in Q_1$ and $q_0\not\in Q_2$. Assume that $\theta  = (r_1>0\wedge\theta_1)\vee(r_2>0\wedge\theta_2)$, $\theta'$ is constructed differently in four cases:
\begin{itemize}
  \item both $I_1$ and $I_2$ do not contain accepting states: $\theta'$ = $\theta $;
  \item $I_1$ contains accepting states while $I_2$ does not: $\theta'$ = $\theta \vee (r_1==0\wedge r_2==0\wedge\theta_1)$;
  \item $I_2$ contains accepting states while $I_1$ does not: $\theta'$ = $\theta \vee (r_1==0\wedge r_2==0\wedge\theta_2)$;
  \item both $I_1$ and $I_2$ contain accepting states: $\theta'$ = $\theta \vee (r_1==0\wedge r_2==0\wedge(\theta_1\vee\theta_2))$;
\end{itemize}

>>>>

\usepackage{standalone}
\usepackage{import}
\usepackage{subcaption}
\usepackage{pgfplots}
\usepackage{color}
\usepackage{tikz}
\usepackage{pgfplots}
\usepgfplotslibrary{fillbetween}
\pgfplotsset{compat=1.16}
\usetikzlibrary {arrows.meta, automata, positioning, fit}
\usepackage{url}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage[]{float}
% set algorithmicx
\renewcommand{\algorithmicrequire}{\textbf{Input:}}  % Use Input in the format of Algorithm  
\renewcommand{\algorithmicensure}{\textbf{Output:}} % Use Output in the format of Algorithm

@@

\newcommand*{\myvec}[1]{\vec{#1}}
\newcommand*{\regex}{\mathcal{R}}
\newcommand*{\strsort}{\verb|Str|}
\newcommand*{\intsort}{\verb|Int|}
\newcommand*{\lan}{\mathcal{L}}
\newcommand*{\highlight}[1]{\textbf{\textit{#1}}}
\newcommand*{\aut}{\mathcal{A}}
\newcommand*{\algfun}[1]{\texttt{#1}}
\newcommand*{\strlen}[1]{\texttt{len}(#1)}
\newcommand*{\myemph}[1]{\textcolor{red}{#1}}
\newcommand*{\mult}[1]{$\times$}
%%
%% end of the preamble, start of the body of the document source.
\begin{document}
%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Efficient Methods for Solving String Constraints with Bounded Repetition}

@@

%% The abstract is a summary of the work to be presented in the
%% article.
\begin{abstract}
\subimport{sections/}{abstract.tex}
\end{abstract}

\section{Introduction} \label{sec:intro}
\subimport{sections/}{introduction.tex}

\section{Preliminaries} \label{sec:pre}
\subimport{sections/}{preliminaries.tex}

\section{Overview} \label{sec:overall}
\subimport{sections/}{overview.tex}

\section{Automaton Model} \label{sec:overview}
\subimport{sections/}{automaton_model.tex}

\section{Solve ESL Formula with Large Repetition Bound} \label{sec:algorithm}
\subimport{sections/}{algorithms.tex}

\section{Implementation and Experimental Results} \label{sec:implementation}
\subimport{sections/}{implementation.tex}

\section{Conclusion and Future Work} \label{sec:conclu}
\subimport{sections/}{conclusion.tex}
